---
title: "Project 4"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(DT)
library(jpeg)
library(ggplot2)
library(tidyverse)
```


```{r global, message=FALSE,warning=FALSE}
# Test if a directory named data exists locally
if (!dir.exists(here::here("data"))) {
    dir.create(here::here("data"))
}

if (!dir.exists(here::here("figures"))) {
    dir.create(here::here("figures"))
}

# saves data only once 
files = c("tuesdata_cocktails.rda")
if (any(!file.exists(here::here("data",files)))) {
    tuesdata <- tidytuesdayR::tt_load('2020-05-26')
    cocktails <- tuesdata$cocktails

    # save the files to RDA objects
    save(cocktails, file = here::here("data","tuesdata_cocktails.rda"))
}

load(here::here("data","tuesdata_cocktails.rda"))

# Optimize the data 
cocktails =cocktails %>%
  mutate(alcoholic = tolower(alcoholic)) %>%
  mutate(alcoholic = if_else(is.na(alcoholic), "Unknown", alcoholic)) %>%
  select(-c(row_id, id_drink, date_modified, iba, video))
```


About {data-icon="ion-information-circled"}
=====================================

**Purpose**

The dashboard allows for audiences to know about different cocktail drinks (alcoholic, non-alcoholic), and also know about the number or name of ingredients used in each drink. This dashboard can help the audiences have a glimpse or evaluation of cocktail drinks that they are intereted in. 

The source showing alcoholic drink data for analysis comes from <a href = "https://www.tidytuesday.com/">TidyTuesday</a>, <a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-05-26">May 26th 2020</a>.

**Video**

<iframe width="1572" height="982" src="https://www.youtube.com/embed/GB1LbKCe3OI"></iframe>


Data {data-icon="ion-android-list"}
=====================================

### Description

The source showing Cocktail data for analysis comes from <a href = "https://www.stephaniehicks.com/jhustatprogramming2023/projects/2023-11-28-project-3/#:~:text=data%20set%20from-,TidyTuesday,-that%20you%20have">TidyTuesday</a>, <a href="https://github.com/rfordatascience/tidytuesday/tree/master/data/2020/2020-05-26">May 26th 2020</a>. 

There are totally 2 different `.csv` data on the website, but I only choose one of them for personal preferences.

To make the data table more convenient to see, I delete some useless columns.

```{r}
renderDataTable({
  data <- head(cocktails, n = 2104)
  DT::datatable(data, options = list(
    bPaginate = FALSE
  ))
})
```



Cocktails Pictures {data-icon="ion-ios-wineglass"}
=====================================

Inputs {.sidebar}
-----------------------------------------------------------------------

```{r}
type = unique(cocktails$alcoholic)
selectInput('type', 'Select Type:', type)


category = unique(cocktails$category)
renderUI({
  selected_type <- input$type
  category_choices <- unique(cocktails$category[cocktails$alcoholic == selected_type])

  selectizeInput("category", "Select Category:", category_choices)
})

drink_choices = unique(cocktails$drink)
renderUI({
  selected_type <- input$type
  selected_category <- input$category
  category_choices <- unique(cocktails$category[cocktails$alcoholic == selected_type])
  
  drink_choices_filtered <- cocktails$drink[cocktails$alcoholic == selected_type & cocktails$category == selected_category]
  
  selectizeInput("drink", "Select Drink:", drink_choices_filtered)
})
```

Row
-----------------------------------------------------------------------

### Data

```{r}
renderTable(
  cocktails %>%
    filter(category == input$category,
           alcoholic == input$type,
           drink == input$drink)
)
```

### Plot {data-height=550}

```{r}
renderPlot({
  max_ing = max(cocktails$ingredient_number)
  min_ing = min(cocktails$ingredient_number)
  selection = cocktails %>%
    filter(category == input$category,
           alcoholic == input$type,
           drink == input$drink) %>%
    summarise(n = n()) %>%
    pull(n)
  
  df = data.frame(max = max_ing,
                  min = min_ing,
                  drink = selection)
  
  df %>%
    ggplot() +
    geom_histogram(aes(x = "Drink Selected", y = drink),
                   stat = "identity", fill = "skyblue") +
    geom_histogram(aes(x = "Max", y = max_ing),
                   stat = "identity", fill = "pink") +
    geom_histogram(aes(x = "Min", y = min_ing),
                   stat = "identity", fill = "lightgreen") +
    coord_flip() +
    labs(title = paste("Number of Ingredients Used in ",input$drink),
         subtitle = "Compared to maximum number of ingredients (12) and minimum number of ingredients (1)",
         x = "Category",
         y = "Count") +
    theme_classic()
  }
)
```



Plot1 {data-icon="ion-android-image"}
=====================================




Analysis {data-icon="ion-ios-analytics"}
=====================================

Row {data-height=450}
-------------------------------------

### Alcoholic Drink Category Count

```{r,comment=NA,fig.width= 7, fig.height=6}
df1 = cocktails %>%
  filter(alcoholic == "alcoholic") %>%
  select(category,drink, ingredient, ingredient_number) %>%
  pivot_wider(names_from = "ingredient_number",
              values_from = "ingredient") %>%
  group_by(category) %>%
  summarize(n = n())

df1 %>%
  ggplot() +
  geom_col(aes(x = category, y = n),
           fill = "#66c2a5") +
  labs(
    title = "Number of Drinks in Alcoholic Cocktail Categories",
    subtitle = "Ordinary Drink has the most count among all alcoholic drink categories\nMilk/Float/Shake has the least count",
    x = "Category",
    y = "Count",
    caption = "Christine Hou"
  ) +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Alcoholic Drink Ingredients Count Distribution

```{r,comment=NA, fig.width=7, fig.height=4}
df2 <- cocktails %>%
  filter(alcoholic == "alcoholic") %>%
  select(drink, category,ingredient, ingredient_number) %>%
  group_by(drink) %>%
  summarize(num_ing = n()) %>%
  group_by(num_ing) %>%
  summarize(num =n()) %>%
  mutate(percentage = num/sum(num)*100) %>%
  mutate(num_ing = as.factor(num_ing))

  
df2 %>%
  ggplot()+
  geom_bar(aes(x = "", y = percentage, fill = num_ing),
           stat = "identity") +
  coord_polar(theta = "y", start=0) +
  theme_void() +
  labs(title = "Distribution of Ingredients Used in Alcoholic Drinks",
    subtitle = "Most alcoholic drinks use 2-5 ingredients",
    caption = "Christine Hou",
    fill = "Number of Ingredients")
```

Row {data-height=550}
-------------------------------------

### Ingredients Distribution Per Alcoholic Drink Category

```{r,comment=NA, fig.width=7, fig.height= 8}
df3 <- cocktails %>%
  filter(alcoholic == "alcoholic") %>%
  select(drink, category,ingredient, ingredient_number) %>%
  group_by(drink,category) %>%
  summarize(num_ing = n(),
            .groups = 'drop')

df3 %>%
  ggplot() +
  geom_boxplot(aes(x = category, y = num_ing)) +
  facet_wrap(~category, scales = "free") +
  theme_minimal() +
  coord_cartesian(ylim = c(0,13)) +
  labs(
    x = "Category",
    y = "Number of Ingredients",
    title = "Number of Ingredients Used in Alcoholic Drinks",
    subtitle = "Drinks belonging to Homemade Liqueur category has the highest number of ingredients (12)\nDrinks belonging to Cocktail and Ordinary Drink category has the least number of ingredients (1)",
    caption = "Christine Hou"
  )
```

